## sf Geometries

```{r, echo = FALSE, include = FALSE}
library(sf)
library(dplyr)
library(magrittr}
```
### Manipulating Geometries

* `st_line_merge`: merges lines
* `st_segmentize`: adds points to straight lines
* `st_voronoi`: creates voronoi tesselation
* `st_centroid`: gives centroid of geometry
* `st_convex_hull`: creates convex hull of set of points
* `st_triangulate`: triangulates set of points (not constrained)
* `st_polygonize`: creates polygon from lines that form a closed ring
* `st_simplify`: simplifies lines by removing vertices
* `st_split`: split a polygon given line geometry
* `st_buffer`: compute a buffer around this geometry/each geometry
* `st_make_valid`: tries to make an invalid geometry valid (requires ``lwgeom``)
* `st_boundary`: return the boundary of a geometry

### convenience functions

* `st_zm`: sets or removes z and/or m geometry
* `st_coordinates`: retrieve coordinates in a matrix or data.frame
* `st_geometry`: set, or retrieve `sfc` from an `sf` object
* `st_is`: check whether geometry is of a particular type

### handling mixes: `GEOMETRY`, `GEOMETRYCOLLECTION`

We can have mixes of geometries with different types at two levels: at the feature level, and at the feature set level. At the feature level:
```{r}
p = st_point(c(0,1))
mp = st_multipoint(rbind(c(1,1), c(3,3)))
ls = st_linestring(rbind(c(1,1), c(3,3)))
(gc = st_geometrycollection(list(p, mp)))
st_sfc(gc)
```

Where do these come from? For instance from `st_intersection`:
```{r}
opar <- par(mfrow = c(1, 2))
a <- st_polygon(list(cbind(c(0,0,7.5,7.5,0),c(0,-1,-1,0,0))))
b <- st_polygon(list(cbind(c(0,1,2,3,4,5,6,7,7,0),c(1,0,.5,0,0,0.5,-0.5,-0.5,1,1))))
plot(a, ylim = c(-1,1))
title("intersecting two polygons:")
plot(b, add = TRUE, border = 'red')
(i <- st_intersection(a,b))
plot(a, ylim = c(-1,1))
title("GEOMETRYCOLLECTION")
plot(b, add = TRUE, border = 'red')
plot(i, add = TRUE, col = 'green', lwd = 2)
par(opar)
```

At the feature set level we end up with `GEOMETRY` objects:
```{r}
(gm = st_sfc(p, mp, ls))
```
How to deal with such messy data? One thing we could do is select:
```{r}
sf = st_sf(a = 1:3, geom = gm)
st_dimension(sf)
sf[st_dimension(sf) == 1,]
```
another is to cast geometries to a common class:
```{r}
st_cast(sf, "MULTIPOINT")
```
which converts the `LINESTRING` to a `MULTIPOINT` as well.

`st_cast` also tries to unravel `GEOMETRYCOLLECTION` geometries in their parts:
```{r}
(parts = st_cast(st_sf(a = 1, geometry = st_sfc(gc))))
```
potentially repeating associated attributes, and tries to find a common class when given no argument 
```{r}
st_cast(parts)
```

### empty geometries

Empty geometries server the role of `NA`, or missing values, for geometries, but they are typed:
```{r}
st_point()
st_geometrycollection()
```

They can be detected by `st_dimension` returning `NA`:
```{r}
st_sfc(st_point(0:1), st_point(), st_geometrycollection()) %>%
	st_dimension
```

```
